Homeless Populations in Toronto

### 2022-02-06 outputs/paper/paper.Rmd
updates

```{r, include = FALSE}
#install.packages("opendatatoronto")
library(opendatatoronto)
library(tidyverse)
library(tidyr)
library(knitr)
library(janitor)
library(lubridate)
library(tibble)
library(dplyr)
library(scales)
```

@@

## Data Characteristics ##

@@

```{r housingattrition, fig.cap = 'Moved to Housing vs Returned from Housing', fig.width = 8, fig.height = 4, echo = FALSE, message = FALSE, warning = FALSE, out.width = '80%', fig.align = "center"}

@@

@@

@@

>>>>

```{r, include = FALSE}
#install.packages("opendatatoronto")
library(opendatatoronto)
library(tidyverse)
library(tidyr)
library(knitr)
library(janitor)
library(lubridate)
library(tibble)
library(dplyr)
library(scales)

```

@@

## Data Analysis ##

@@

```{r housingattrition, fig.cap = 'Moved to Housing vs Returned from Housing', fig.width = 8, fig.height = 2.5, echo = FALSE, fig.align = "center", fig.pos = 'h'}

@@

The next aspect of the data I analyzed was the composition of the total homeless population and its fluctuations over the last two years. If applicable, the Shelter Intake Management System further classifies homeless people into the following sub-groups:

  * Chronic: Refers to chronic homelessness, which is defined as someone who has recorded a minimum of 180 overnight stays in the              past year, or has recorded a cumulative 546 overnight stays in the past 3 years.

  * Families: Individuals staying in a family designated overnight service

  * Youth: Individuals that are between 16 and 24 years of age by the end of the reporting month, and who are not members of a               family, as defined above.

  * Single Adult: Individuals that are neither classified as Youth nor as part of a family, as defined above.

  * Refugees: Individuals that either identify as refugees when registering with a shelter system, or who register into a program               designated for refugees. When registering a family, all members of the family are registered in the same way as the               household head.

  * Non-Refugees: People who are not classified as refugees.

  * Indigenous: A person who has self-identified as First Nations, Metis or Inuit in at least one of the intakes completed by SMIS

Identifying the which groups are most prevalent in shelters can allow for better allocation of funds in order to maximize the benefits to the populations most in need. Note that people only started being classified as Indigenous starting in October 2020, and it took a brief period for the data to stabilize. As a result, data on Indigenous people only began to be included on January 2021. Figure \@ref(fig:popcomposition) shows the proportion of the total homeless population accounted for by these sub-groups. Here we can see that since January 2020, the proportion of non-refugees has increased from 70% to almost 85%, while the proportion of chronically homeless people has increased by about 10%, from 35% to over 45%. Notably, the youth population has stayed quite consistent at about 10% of the total homeless population, while the proportion of single adults has increased from about 63% to over 70%.

```{r popcomposition, fig.cap = 'Total Population Composition Since 2020', fig.width = 6, fig.height = 3, echo = FALSE, fig.align = "center", fig.pos = 'h'}
homeless_comp <-
  populations %>% 
  filter(population_group != "All Population") %>% 
  select(population_group, Date, actively_homeless, population_group_percentage)

homeless_comp %>% 
  ggplot(aes(x = Date, y = population_group_percentage, color = population_group, group = 2)) +
  geom_point() +
  geom_line(aes(group = population_group)) +
  labs(title = "Total Population Composition Over Since 2020") +
  ylab("Percentage of Total Population") +
  scale_y_continuous(breaks = pretty(homeless_comp$population_group_percentage, n = 10))
```

@@

jabbadoo

@@

bazoo
